---
title: "meta_BPA_cancer_remove"
author: "Yu He"
date: "2024-03-07"
output: 
  html_document:
    fig_width: 12
    fig_height: 10
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(meta)
library(readxl)
library(metafor)
```


# Data Import，remove 'Keshavarz4_urine', 'Keshavarz4_tissue', 'Hong8', 'Marotta23' study due to large 95% CI
```{r}
setwd("~/Desktop/meta analysis/BPA_cancer")

# load the data with strongest effect of OR
meta_data_s_remove <- read_excel("data.xlsx", sheet = "strong_or") %>% 
  janitor::clean_names() %>% 
  filter(!(auther %in% c('Keshavarz4_urine', 'Keshavarz4_tissue', 'Hong8', 'Marotta23'))) 

# load the data with average effect of OR
meta_data_a_remove <- read_excel("data.xlsx", sheet = "average_or") %>% 
  janitor::clean_names() %>% 
  filter(!(auther %in% c('Keshavarz4_urine', 'Keshavarz4_tissue', 'Hong8', 'Marotta23'))) 

# Calculate the log of odds ratios and their standard errors
# strong effect of OR
meta_data_s_remove <- meta_data_s_remove %>% 
  mutate(
    log_odds_ratio = log(or),
    lower_log =log(lower_ci),
    upper_log = log(upper_ci),
    se_log_odds_ratio = (upper_log - lower_log) / (2 * 1.96) # 1.96 is the Z value for 95% CI
  )

# average OR
meta_data_a_remove <-meta_data_a_remove %>% 
  mutate(
    log_odds_ratio = log(or),
    lower_log =log(lower_ci),
    upper_log = log(upper_ci),
    se_log_odds_ratio = (upper_log - lower_log) / (2 * 1.96) # 1.96 is the Z value for 95% CI
  )
```


# Part 1, general meta analysis
```{r}
# strong effect OR
meta_s_analysis_remove <- metagen(
  TE = log_odds_ratio,
  seTE = se_log_odds_ratio,
  studlab = paste(auther),
  data = meta_data_s_remove,
  sm = "OR"
)

summary(meta_s_analysis_remove)

# average OR
meta_a_analysis_remove <- metagen(
  TE = log_odds_ratio,
  seTE = se_log_odds_ratio,
  studlab = paste(auther),
  data = meta_data_a_remove,
  sm = "OR"
)

summary(meta_a_analysis_remove)


# Forest plot
# strong effect of OR
# pdf(file = "forestplot_strongeffect.pdf", width = 12, height = 10), use to save the forest plot
forest(meta_s_analysis_remove, 
       sortvar = TE,
       prediction = TRUE,
       leftlabs = c("Author", "g", "SE")
       )

# average OR
# pdf(file = "forestplot_average.pdf", width = 12, height = 10)
forest(meta_a_analysis_remove, 
       sortvar = TE,
       prediction = TRUE,
       leftlabs = c("Author", "g", "SE")
       )
# dev.off(), use to save the forest plot
```


## Assess the publication bias of general meta analysis
```{r}
# Funnel plot
# strong effect of OR
# pdf(file = "funnelplot_strongeffect.pdf", width = 12, height = 10)
funnel(meta_s_analysis_remove)
# average OR
# pdf(file = "funnelplot_average.pdf", width = 12, height = 10)
funnel(meta_a_analysis_remove)
# dev.off()

# assess the asymmetry of funnel plot and publication bias
# Conduct the meta-analysis using rma() function from metafor
# strong effect of OR
meta_s_analysis_rma <- rma(yi = log_odds_ratio, sei = se_log_odds_ratio, method = "REML", data = meta_data_s_remove)

# Perform Egger's test for publication bias
eggers_test_s <- regtest(meta_s_analysis_rma, model = "lm")

# Output the results of Egger's test
summary(meta_s_analysis_rma)
summary(eggers_test_s$fit)

# average OR
meta_a_analysis_rma <- rma(yi = log_odds_ratio, sei = se_log_odds_ratio, method = "REML", data = meta_data_a_remove)

# Perform Egger's test for publication bias
eggers_test_a <- regtest(meta_a_analysis_rma, model = "lm")

# Output the results of Egger's test
summary(meta_a_analysis_rma)
summary(eggers_test_a$fit)
```


```{r}
# Begg's Test
# strong effect of OR
begg_test_s <- ranktest(meta_s_analysis_rma)
print(begg_test_s)

# average OR
begg_test_a <- ranktest(meta_a_analysis_rma)
print(begg_test_a)
```


```{r}
# Trim and Fill Method
# strong effect of OR
trimfill_result_s <- trimfill(meta_s_analysis_rma, estimator = "R0")
print(trimfill_result_s)

# average OR
trimfill_result_a <- trimfill(meta_a_analysis_rma, estimator = "R0")
print(trimfill_result_a)
```




# Part 2, subgroup region, China vs None-China
```{r}
# subgroup - nation 
# strong effect OR
meta_data_s_na <- meta_data_s_remove %>% 
  mutate(
    region_code = ifelse(region == "China", 0, 1)
  )

meta_s_analysis_china <- metagen(
  TE = log_odds_ratio,
  seTE = se_log_odds_ratio,
  studlab = paste(auther),
  data = subset(meta_data_s_na, region_code == "0"), # China
  sm = "OR"
)


meta_s_analysis_other <- metagen(
  TE = log_odds_ratio,
  seTE = se_log_odds_ratio,
  studlab = paste(auther),
  data = subset(meta_data_s_na, region_code == "1"), # other region
  sm = "OR"
)

summary(meta_s_analysis_china)
summary(meta_s_analysis_other)
summary(meta_s_analysis_remove)

# average OR
meta_data_a_na <- meta_data_a_remove %>% 
  mutate(
    region_code = ifelse(region == "China", 0, 1)
  )

meta_a_analysis_china <- metagen(
  TE = log_odds_ratio,
  seTE = se_log_odds_ratio,
  studlab = paste(auther),
  data = subset(meta_data_a_na, region_code == "0"), # China
  sm = "OR"
)

meta_a_analysis_other <- metagen(
  TE = log_odds_ratio,
  seTE = se_log_odds_ratio,
  studlab = paste(auther),
  data = subset(meta_data_a_na, region_code == "1"), # other region
  sm = "OR"
)

summary(meta_a_analysis_china)
summary(meta_a_analysis_other)
summary(meta_a_analysis_remove)
```



```{r}
# Forest plot
# strong effect of OR
#pdf(file = "forestplot_strongeffect_china.pdf", width = 12, height = 10)
forest(meta_s_analysis_china, 
       sortvar = TE,
       prediction = TRUE,
       leftlabs = c("Author", "g", "SE")
       )

#pdf(file = "forestplot_strongeffect_otherregion.pdf", width = 12, height = 10)
forest(meta_s_analysis_other, 
       sortvar = TE,
       prediction = TRUE,
       leftlabs = c("Author", "g", "SE")
       )

# average OR
#pdf(file = "forestplot_average_china.pdf", width = 12, height = 10)
forest(meta_a_analysis_china, 
       sortvar = TE,
       prediction = TRUE,
       leftlabs = c("Author", "g", "SE")
       )


#pdf(file = "forestplot_average_otherregion.pdf", width = 12, height = 10)
forest(meta_a_analysis_other, 
       sortvar = TE,
       prediction = TRUE,
       leftlabs = c("Author", "g", "SE")
       )

#dev.off()
```



## Assess the publication bias of subgroup - China
```{r}
# Funnel plot
# strong effect of OR
# pdf(file = "funnelplot_strongeffect_china.pdf", width = 12, height = 10)
funnel(meta_s_analysis_china)
# pdf(file = "funnelplot_strongeffect_otherregion.pdf", width = 12, height = 10)
funnel(meta_s_analysis_other)
# average OR
# pdf(file = "funnelplot_average_china.pdf", width = 12, height = 10)
funnel(meta_a_analysis_china)
# pdf(file = "funnelplot_average_otherregion.pdf", width = 12, height = 10)
funnel(meta_a_analysis_other)
# dev.off()

```


```{r}
# assess the asymmetry of funnel plot and publication bias
# Conduct the meta-analysis using rma() function from metafor
# strong effect of OR
meta_s_analysis_rma_china <- rma(yi = log_odds_ratio, sei = se_log_odds_ratio, method = "REML", data = subset(meta_data_s_na, region_code == "0"))

# Perform Egger's test for publication bias
eggers_test_s_china <- regtest(meta_s_analysis_rma_china, model = "lm")

# Output the results of Egger's test
summary(eggers_test_s_china$fit)

# average OR
meta_a_analysis_rma_china <- rma(yi = log_odds_ratio, sei = se_log_odds_ratio, method = "REML", data = subset(meta_data_a_na, region_code == "0"))

# Perform Egger's test for publication bias
eggers_test_a_china <- regtest(meta_a_analysis_rma_china, model = "lm")

# Output the results of Egger's test
summary(eggers_test_a_china$fit)
```


```{r}
# Begg's Test
# strong effect of OR
begg_test_s_china <- ranktest(meta_s_analysis_rma_china)
print(begg_test_s_china)

# average OR
begg_test_a_china <- ranktest(meta_a_analysis_rma_china)
print(begg_test_a_china)
```


```{r}
# Trim and Fill Method
# strong effect of OR
trimfill_result_s_china <- trimfill(meta_s_analysis_rma_china, estimator = "R0")
summary(trimfill_result_s_china)

# average OR
trimfill_result_a_china <- trimfill(meta_a_analysis_rma_china, estimator = "R0")
summary(trimfill_result_a_china)
```


## Assess the publication bias of subgroup - other region
```{r}
# assess the asymmetry of funnel plot and publication bias
# Conduct the meta-analysis using rma() function from metafor
# strong effect of OR
meta_s_analysis_rma_other <- rma(yi = log_odds_ratio, sei = se_log_odds_ratio, method = "REML", data = subset(meta_data_a_na, region_code == "1"))

# Perform Egger's test for publication bias
eggers_test_s_other <- regtest(meta_s_analysis_rma_other, model = "lm")

# Output the results of Egger's test
summary(eggers_test_s_other$fit)

# average OR
meta_a_analysis_rma_other <- rma(yi = log_odds_ratio, sei = se_log_odds_ratio, method = "REML", data = subset(meta_data_a_na, region_code == "0"))

# Perform Egger's test for publication bias
eggers_test_a_other <- regtest(meta_a_analysis_rma_other, model = "lm")

# Output the results of Egger's test
summary(eggers_test_a_other$fit)
```


```{r}
# Begg's Test
# strong effect of OR
begg_test_s_other <- ranktest(meta_s_analysis_rma_other)
begg_test_s_other

# average OR
begg_test_a_other <- ranktest(meta_a_analysis_rma_other)
begg_test_a_other
```


```{r}
# Trim and Fill Method
# strong effect of OR
trimfill_result_s_other <- trimfill(meta_s_analysis_rma_other, estimator = "R0")
trimfill_result_s_other

# average OR
trimfill_result_a_other <- trimfill(meta_a_analysis_rma_other, estimator = "R0")
trimfill_result_a_other
```


# Part 3, subgroup study design, cross sectional vs case control
```{r}
# re-code the data
# strong effect of OR
meta_data_s_sen <- meta_data_s_remove %>% 
  mutate(
    study_design = factor(study_design, levels = c(0, 1), labels = c("case-control", "cross-sectional"))
  )

# average OR
meta_data_a_sen <- meta_data_a_remove %>% 
  mutate(
    study_design = factor(study_design, levels = c(0, 1), labels = c("case-control", "cross-sectional"))
  )
```


```{r}
# strong effect of OR 
# case control study
meta_s_analysis_cc <- metagen(
  TE = log_odds_ratio,
  seTE = se_log_odds_ratio,
  studlab = paste(auther),
  data = subset(meta_data_s_sen, study_design == "case-control"),
  sm = "OR"
)

# cross sectional study
meta_s_analysis_cs <- metagen(
  TE = log_odds_ratio,
  seTE = se_log_odds_ratio,
  studlab = paste(auther),
  data = subset(meta_data_s_sen, study_design == "cross-sectional"),
  sm = "OR"
)

# average OR 
# case control study
meta_a_analysis_cc <- metagen(
  TE = log_odds_ratio,
  seTE = se_log_odds_ratio,
  studlab = paste(auther),
  data = subset(meta_data_a_sen, study_design == "case-control"),
  sm = "OR"
)

# cross sectional study
meta_a_analysis_cs <- metagen(
  TE = log_odds_ratio,
  seTE = se_log_odds_ratio,
  studlab = paste(auther),
  data = subset(meta_data_a_sen, study_design == "cross-sectional"),
  sm = "OR"
)

# Summary of the meta-analysis
# strong effect of OR
summary(meta_s_analysis_cc)
summary(meta_s_analysis_cs)
summary(meta_s_analysis_remove)

# average OR
summary(meta_a_analysis_cc)
summary(meta_a_analysis_cs)
summary(meta_a_analysis_remove)
```


```{r}
# Forest plot
# strong effect of OR
#pdf(file = "forestplot_strongeffect_cc.pdf", width = 12, height = 10)
forest(meta_s_analysis_cc, 
       sortvar = TE,
       prediction = TRUE,
       leftlabs = c("Author", "g", "SE")
       )

#pdf(file = "forestplot_strongeffect_cs.pdf", width = 12, height = 10)
forest(meta_s_analysis_cs, 
       sortvar = TE,
       prediction = TRUE,
       leftlabs = c("Author", "g", "SE")
       )

# average OR
#pdf(file = "forestplot_average_cc.pdf", width = 12, height = 10)
forest(meta_a_analysis_cc, 
       sortvar = TE,
       prediction = TRUE,
       leftlabs = c("Author", "g", "SE")
       )


#pdf(file = "forestplot_average_cs.pdf", width = 12, height = 10)
forest(meta_a_analysis_cs, 
       sortvar = TE,
       prediction = TRUE,
       leftlabs = c("Author", "g", "SE")
       )

#dev.off()
```


```{r}
# Funnel plot 
# strong effect of OR
#pdf(file = "fuunelplot_strongeffect_cc.pdf", width = 12, height = 10)
funnel(meta_s_analysis_cc)
#pdf(file = "fuunelplot_strongeffect_cs.pdf", width = 12, height = 10)
funnel(meta_s_analysis_cs)
# average OR
#pdf(file = "fuunelplot_average_cc.pdf", width = 12, height = 10)
funnel(meta_a_analysis_cc)
#pdf(file = "fuunelplot_average_cs.pdf", width = 12, height = 10)
funnel(meta_s_analysis_cs)
#dev.off()
```


## Assess the publication bias subgroup study design, case control
```{r}
# assess the asymmetry of funnel plot and publication bias
# Conduct the meta-analysis using rma() function from metafor
# strong effect of OR
meta_s_analysis_rma_cc <- rma(yi = log_odds_ratio, sei = se_log_odds_ratio, method = "REML", data = subset(meta_data_s_sen, study_design == "case-control"))

# Perform Egger's test for publication bias
eggers_test_s_cc <- regtest(meta_s_analysis_rma_cc, model = "lm")

# Output the results of Egger's test
summary(eggers_test_s_cc$fit)

# average OR
meta_a_analysis_rma_cc <- rma(yi = log_odds_ratio, sei = se_log_odds_ratio, method = "REML", data = subset(meta_data_a_sen, study_design == "case-control"))

# Perform Egger's test for publication bias
eggers_test_a_cc <- regtest(meta_a_analysis_rma_cc, model = "lm")

# Output the results of Egger's test
summary(eggers_test_a_cc$fit)
```


```{r}
# Begg's Test
# strong effect of OR
begg_test_s_cc <- ranktest(meta_s_analysis_rma_cc)
print(begg_test_s_cc)

# average OR
begg_test_a_cc <- ranktest(meta_a_analysis_rma_cc)
print(begg_test_a_cc)
```


```{r}
# Trim and Fill Method
# strong effect of OR
trimfill_result_s_cc <- trimfill(meta_s_analysis_rma_cc, estimator = "R0")
print(trimfill_result_s_cc)

# average OR
trimfill_result_a_cc <- trimfill(meta_a_analysis_rma_cc, estimator = "R0")
print(trimfill_result_a_cc)
```


## Assess the publication bias subgroup study design, cross sectional
```{r}
# assess the asymmetry of funnel plot and publication bias
# Conduct the meta-analysis using rma() function from metafor
# strong effect of OR
meta_s_analysis_rma_cs <- rma(yi = log_odds_ratio, sei = se_log_odds_ratio, method = "REML", data = subset(meta_data_s_sen, study_design == "cross-sectional"))

# Perform Egger's test for publication bias
eggers_test_s_cs <- regtest(meta_s_analysis_rma_cs, model = "lm")

# Output the results of Egger's test
summary(eggers_test_s_cs$fit)

# average OR
meta_a_analysis_rma_cs <- rma(yi = log_odds_ratio, sei = se_log_odds_ratio, method = "REML", data = subset(meta_data_a_sen, study_design == "cross-sectional"))

# Perform Egger's test for publication bias
eggers_test_a_cs <- regtest(meta_a_analysis_rma_cs, model = "lm")

# Output the results of Egger's test
summary(eggers_test_a_cs$fit)
```


```{r}
# Begg's Test
# strong effect of OR
begg_test_s_cs <- ranktest(meta_s_analysis_rma_cs)
print(begg_test_s_cs)

# average OR
begg_test_a_cs <- ranktest(meta_a_analysis_rma_cs)
print(begg_test_a_cs)
```


```{r}
# Trim and Fill Method
# strong effect of OR
trimfill_result_s_cs <- trimfill(meta_s_analysis_rma_cs, estimator = "R0")
print(trimfill_result_s_cs)

# average OR
trimfill_result_a_cs <- trimfill(meta_a_analysis_rma_cs, estimator = "R0")
print(trimfill_result_a_cs)
```


# Part 4, subgroup sample method, urine (= 1) vs other (= 0, 2, and 3)
```{r}
# re-code the data
# strong effect of OR
meta_data_s_sm <- meta_data_s_remove %>% 
  mutate(
    samplemethod_code = ifelse(sample_method %in% c("1", "2"), 1, 0) # urine = 1, other = 0
  ) 

# average OR
meta_data_a_sm <- meta_data_a_remove %>% 
  mutate(
    samplemethod_code = ifelse(sample_method %in% c("1", "2"), 1, 0) # urine = 1, other = 0
  )
```


```{r}
# strong effect of OR, urine = 1, other = 0
# urine sample method
meta_s_analysis_ur <- metagen(
  TE = log_odds_ratio,
  seTE = se_log_odds_ratio,
  studlab = paste(auther),
  data = subset(meta_data_s_sm, samplemethod_code == "1"),
  sm = "OR"
)

# other sample method
meta_s_analysis_ot <- metagen(
  TE = log_odds_ratio,
  seTE = se_log_odds_ratio,
  studlab = paste(auther),
  data = subset(meta_data_s_sm, samplemethod_code == "0"),
  sm = "OR"
)

# average OR, urine = 1, other = 0 
# urine sample method
meta_a_analysis_ur <- metagen(
  TE = log_odds_ratio,
  seTE = se_log_odds_ratio,
  studlab = paste(auther),
  data = subset(meta_data_a_sm, samplemethod_code == "1"),
  sm = "OR"
)

# other sample method
meta_a_analysis_ot <- metagen(
  TE = log_odds_ratio,
  seTE = se_log_odds_ratio,
  studlab = paste(auther),
  data = subset(meta_data_a_sm, samplemethod_code == "0"),
  sm = "OR"
)

# Summary of the meta-analysis
# strong effect of OR
summary(meta_s_analysis_ur)
summary(meta_s_analysis_ot)
summary(meta_s_analysis_remove)

# average OR
summary(meta_a_analysis_ur)
summary(meta_a_analysis_ot)
summary(meta_a_analysis_remove)
```


```{r}
# Forest plot
# strong effect of OR
#pdf(file = "forestplot_strongeffect_ur.pdf", width = 12, height = 10)
forest(meta_s_analysis_ur, 
       sortvar = TE,
       prediction = TRUE,
       leftlabs = c("Author", "g", "SE")
       )

#pdf(file = "forestplot_strongeffect_ot.pdf", width = 12, height = 10)
forest(meta_s_analysis_ot, 
       sortvar = TE,
       prediction = TRUE,
       leftlabs = c("Author", "g", "SE")
       )

# average OR
#pdf(file = "forestplot_average_ur.pdf", width = 12, height = 10)
forest(meta_a_analysis_ur, 
       sortvar = TE,
       prediction = TRUE,
       leftlabs = c("Author", "g", "SE")
       )


#pdf(file = "forestplot_average_ot.pdf", width = 12, height = 10)
forest(meta_a_analysis_ot, 
       sortvar = TE,
       prediction = TRUE,
       leftlabs = c("Author", "g", "SE")
       )

#dev.off()
```


```{r}
# Funnel plot 
# strong effect of OR
#pdf(file = "fuunelplot_strongeffect_ur.pdf", width = 12, height = 10)
funnel(meta_s_analysis_ur)
#pdf(file = "fuunelplot_strongeffect_ot.pdf", width = 12, height = 10)
funnel(meta_s_analysis_ot)
# average OR
#pdf(file = "fuunelplot_average_ur.pdf", width = 12, height = 10)
funnel(meta_a_analysis_ur)
#pdf(file = "fuunelplot_average_ot.pdf", width = 12, height = 10)
funnel(meta_s_analysis_ot)
#dev.off()
```

## Assess the publication bias, subgroup sample method, urine (= 1) 
```{r}
# assess the asymmetry of funnel plot and publication bias
# Conduct the meta-analysis using rma() function from metafor
# strong effect of OR
meta_s_analysis_rma_ur <- rma(yi = log_odds_ratio, sei = se_log_odds_ratio, method = "REML", data = subset(meta_data_s_sm, samplemethod_code == "1"))

# Perform Egger's test for publication bias
eggers_test_s_ur <- regtest(meta_s_analysis_rma_ur, model = "lm")

# Output the results of Egger's test
summary(eggers_test_s_ur$fit)

# average OR
meta_a_analysis_rma_ur <- rma(yi = log_odds_ratio, sei = se_log_odds_ratio, method = "REML", data = subset(meta_data_a_sm, samplemethod_code == "1"))

# Perform Egger's test for publication bias
eggers_test_a_ur <- regtest(meta_a_analysis_rma_ur, model = "lm")

# Output the results of Egger's test
summary(eggers_test_a_ur$fit)
```


```{r}
# Begg's Test
# strong effect of OR
begg_test_s_ur <- ranktest(meta_s_analysis_rma_ur)
print(begg_test_s_ur)

# average OR
begg_test_a_ur <- ranktest(meta_a_analysis_rma_ur)
print(begg_test_a_ur)
```


```{r}
# Trim and Fill Method
# strong effect of OR
trimfill_result_s_ur <- trimfill(meta_s_analysis_rma_ur, estimator = "R0")
print(trimfill_result_s_ur)

# average OR
trimfill_result_a_ur <- trimfill(meta_a_analysis_rma_ur, estimator = "R0")
print(trimfill_result_a_ur)
```


## Assess the publication bias, subgroup sample method, other (= 0, 2, and 3)
```{r}
# assess the asymmetry of funnel plot and publication bias
# Conduct the meta-analysis using rma() function from metafor
# strong effect of OR
meta_s_analysis_rma_ot <- rma(yi = log_odds_ratio, sei = se_log_odds_ratio, method = "REML", data = subset(meta_data_s_sm, samplemethod_code == "0"))

# Perform Egger's test for publication bias
eggers_test_s_ot <- regtest(meta_s_analysis_rma_ot, model = "lm")

# Output the results of Egger's test
summary(eggers_test_s_ot$fit)

# average OR
meta_a_analysis_rma_ot <- rma(yi = log_odds_ratio, sei = se_log_odds_ratio, method = "REML", data = subset(meta_data_a_sm, samplemethod_code == "0"))

# Perform Egger's test for publication bias
eggers_test_a_ot <- regtest(meta_a_analysis_rma_ot, model = "lm")

# Output the results of Egger's test
summary(eggers_test_a_ot$fit)
```


```{r}
# Begg's Test
# strong effect of OR
begg_test_s_ot <- ranktest(meta_s_analysis_rma_ot)
print(begg_test_s_ot)

# average OR
begg_test_a_ot <- ranktest(meta_a_analysis_rma_ot)
print(begg_test_a_ot)
```


```{r}
# Trim and Fill Method
# strong effect of OR
trimfill_result_s_ot <- trimfill(meta_s_analysis_rma_ot, estimator = "R0")
print(trimfill_result_s_ot)

# average OR
trimfill_result_a_ot <- trimfill(meta_a_analysis_rma_ot, estimator = "R0")
print(trimfill_result_a_ot)
```











